\cleartooddpage[\thispagestyle{empty}]
\chapter{Gamma Rays}


\section{Methods of Creation}

\subsection{Leptonic Processes}
Electrons up-scatter ambient photons into higher-energy gamma rays.
Need an electron accelerator, and need photons of a necessary spectrum.
Spinning pulsars, when their magnetic field lines break and reconnect, can accelerate electrons. (??)

\subsection{Hadronic Processes}
High energy protons collide with other protons, decay into pions, and the neutral pions ($\pi_{0}$) from that decay into gamma-ray pairs.
These protons can be locally accelerated by a nearby mass of either stellar-scale (example??) or galaxy-scale(example??), or may be from other external galaxies.

\subsection{Dark Matter Interactions}
Indirect Detection: dark-dark => sm-sm
Direct Detection  : dark-sm => dark-sm
Collider Detection: sm-sm => dark-dark

Indirect can produce gamma rays via quark/lepton annihilation (??)
dark-dark => t-t => gamma-gamma
dark-dark => b-b => gamma-gamma


\section{Atmospheric Shower}
(this entire section needs more citations!??)

When a gamma ray, proton, or other particle strikes an atom of Earth's atmosphere, it can set off a cascade of energetic particles called an air shower.
When the primary particle is a gamma ray, an electron, or a positron, it creates an electromagnetic shower.
When the primary is a proton or other baryon, it creates a hadronic shower.
During either cascade of particles, any charged particles travelling at $v_{particle} > c_{atmosphere}$ will create Cherenkov photons, UV- and Visible-spectrum photons that are then imaged and recorded by the VERITAS observatory.

(image of particle cascade diagram??)

Electromagnetic air showers are started by a high energy (\nicetilde TeV) electron or gamma-ray, and produce a cascade of electrons, positrons, and photons, where initially each successive generation of particles tends to have more particles and less energy per particle than the last.
The primary gamma ray will interact with an atmospheric atom, producing a $e^{-}e^{+}$ pair, each with roughly half the primary gamma ray's energy.
The $e^{-}$ and $e^{+}$ will emit some photons through bremstrahlung radiation, until they only have a few MeV of kinetic energy, after which other energy loss mechanisms (compton, etc) dominate.
The photons created during the shower go on to produce more $e^{-}e^{+}$ pairs, though as each newly created particle has less energy than its parent particle, eventually the photons in the shower don't have enough energy to produce additional $e^{-}e^{+}$ pairs, and the shower dies off.

As most (\nicetilde 99\%??) of detected air showers are due to protons and not gamma rays, understanding the differences between hadronic showers and electromagnetic showers becomes useful in removing unwanted proton air showers and preserving gamma-ray air showers within the reconstruction software, sometimes referred to as gamma-hadron separation.
Hadronic showers start with a primary \nicetilde TeV proton that interacts with an atmospheric atom.
This proton then converts into a $\pi^{+}\pi^{-}\pi^{0}$, each with roughly \nicetilde 33\% (it varies though??) of the initial proton's energy.
The $\pi^{+}$ and $\pi^{-}$ can travel far from the main axis of the primary particle, then produce $\mu\nu$ pairs.
The $\pi^{0}$ quickly decays into $\gamma\gamma$, which then each start their own electromagnetic shower.
The $\pi^{+}$ and $\pi^{-}$ can have a large transverse momentum (relative to the primary shower axis), and also have longer decay time (??), both of which contribute to creating sub-cascades of showers further away from the primary particle axis, which tends to cause hadronic showers (and their resulting Cherenkov images) to be wider than a purely electromagnetic shower of the same length. 

(image of proton vs gamma shower??)

(image of proton vs gamma shower cherenkov image??)


\section{Galactic Center}
The galactic center is a complex region of space, with many astrophysical sources of gamma rays.
A disk of dust lies along the galactic plane, acting as an interaction medium for diffuse proton cosmic rays.
Nearby supernova remnants also produce gamma-rays as their expanding shell interacts with ambient dust.
The immediate area surrounding the galactic center a point-source emitter of gamma rays, though this mechanism is uncertain.

\subsection{Supermassive Black Hole}

Through kinematic observations of nearby stars, the galactic center is suspected to have a black hole, on the order of $10^6 M_{\odot}$. ??

\subsection{Diffuse Emission}
The disk of gas that permeates the galactic plane acts as an interaction medium for passing cosmic rays, both from galactic accelerators, as well as from extragalactic sources.
These high-energy protons collide with the dust, shattering into $\pi^{\pm,0}$.
The $\pi^0$ then decays into two gamma rays ??, providing the diffuse emission.
The spectrum of this emission is ??.


