\cleartooddpage[\thispagestyle{empty}]
\chapter{Introduction}

How much mass is contained within a given volume of space?
A Galaxy?
A Cluster of Galaxies?
The observable universe?

Questions of 'How Much Mass is in a Region of Space' have been asked since the 1930s, but different methods of measuring the answer always disagree.
Observing cosmological and gravitational features results in a larger answer, while observing the quantity of light produced results in a smaller one.
This implies that there is missing mass, missing matter, unaccounted for by our current models of the universe.
As this matter prefers to only interact gravitationally, without interacting with the electromagnetic spectrum, it has earned the (in)conspicuous title of Dark Matter.


\section{Dark Matter}

Dark matter is measured with several methods, described here in order of the volume of space measured.

At the smallest scale, satellite galaxies consist of several thousand, individually-resolvable stars.
Telescopes can measure the spectra of these stars, allowing for their velocity to be calculated.
By looking at how fast each star is moving around the satellite galaxy's center of mass, the amount of enclosed mass can be inferred.
??Volume measured??

At the next highest scale, the total amount of light produced by a quadrant of a galaxy is measured with telescopes.
Known mass-to-light ratios can then be used to calculate the total amount of mass within that quadrant.
This was the earliest method of measuring the mass of a galaxy, producing a low calculation for the amount of mass.
??Volume measured??

In the third method, a galaxy's emission spectrum is observed at many positions around its structure (center, outer edges, etc).
By comparing the orientation of the disk with the doppler-shifted position of common spectral lines, one can calculate the average velocity that each chunk is travelling at around the center of the galaxy.
Newton's law of gravity can then be used to calculate the mass contained within a sphere of that same radius.
This calculation ends up with a larger amount of mass.

In the fourth method, a large swath of space can be observed with high-resolution telescopes.
Volumes of space with large amounts of matter will bend the light from background galaxies in known ways.
By examining how the images of different background galaxies are bent, the amount of mass between the galaxy and the observer can be inferred.
This is typically used for measuring the mass of cluster-sized volumes of space, rather than individual galaxies.

At the largest scale, the cosmic microwave background (CMB) has been used to measure the total amount of dark matter in the universe.
By looking at the structure of the CMB, earlier times in the universe can be studied.
Closer to the big bang, there was a point in time referred to as the 'freezout'.
Before that time, the universe was a sea of quark plasma and other particles, and the temperature was too hot to allow baryons exist for any long period of time.
After that freezout time, the universe had expanded enough, and the temperature had dropped to the point where more baryons were created than destroyed, much like cooling steam condensing to liquid water.
This freezout left an imprint of its structure, more CMB photons in some places, less in others.
By looking at this CMB structure, the total number of baryons in the universe can be measured.
??Volume measured??




\section{Galactic Center}

At the center of every galaxy there is a supermassive black hole, with a mass of millions of suns.
The Milky Way's own supermassive black hole is measured to be about $4*10^6$ solar masses.

\section{Gamma-Ray Astronomy}
Humans have been attempting to detect gamma rays for several decades now, and they're starting to get the hang of it.
Usually this is done by chosing a volume of solid, liquid, or gas matter, setting up sensitive particle detectors, waiting for gamma rays to strike the mass, and observing the particles that are emitted by the strike.
As even the the massive emitted particles are still travelling at a large fraction of light-speed, very high-speed electronics are needed to record the signals.

\section{Science Objectives}
The goal of this thesis is to demonstrate that, by analyzing gamma-rays from the galactic center, a search for a Dark Matter Halo can be conducted.
If a halo is discovered, it would be a huge discovery.
If a halo is not discovered, this allows for upper limits to be placed on the halo's flux.

\section{Gamma-Ray Showers}
Gamma rays are very high energy photons.
When these photons strike the Earth, the majority spallate off of atoms in the atmosphere.
This spallation consumes the photo to produce an electron-positron pair, which then each strike other particles in the atmosphere.
The electrons tend to strike other atomic electrons, knocking them free, while the positrons will annihilate with other atomic electrons, producing a pair of photons.
These particles continue to strike and divide, cascading in an oval-shaped shower of energetic particles along the original direction of the primary gamma ray.
Eventually, as the number of particles in each successive generation is more numerous than the previous, the average energy of each particle goes down.
Once the electrons fall to around 80 MeV ??, then they lose energy rapidly to ionization of atmospheric atoms, and are no longer able to create new generations of particles.
Throughout this shower, the charged electrons and positrons are passing through the atmosphere at a speed very close to $c_{vaccume}$.
As the speed of light in the atmosphere is less than the relativistic speeds of the particles, the charged particles will travel through the atmosphere, outracing the effects of their charge on nearby atmospheric molecules.
This outracing creates a wave of polarziation in the atmospheric molecules, resulting in the emission of lower-energy (visible and ultraviolet) photons known as Cherenkov photons, at a small angle relative to the charged particles' direction of travel.

\section{Gamma Ray Observatories}
Modern gamma-ray observatories usually function by detecting the particles created by the gamma ray as it strikes some known interaction mass.
The interaction mass can be as small as 1 $m^3$ or as large as the atmosphere on earth.
Veritas functions by watching for gamma-ray showers, by using Earth's atmosphere as the interaction mass.
Gamma-ray showers produce Cherenkov photons in a very short amount of time, typically on the order of ~10s of nanoseconds.
By constructing a specialized camera that counts the Cherenkov photons from the sky at ~1ns intervals, the gamma ray's shower can be imaged.
By combining multiple synchronized cameras spaced at roughly ~100m apart, multiple images can be taken of the same showers.
As the showers are highly directional along the original gamma-ray's trajectory, the images then point backwards to the original gamma-ray's point-of-origin in the sky.
By running a large amount of simulations, a database of shower sizes and positions can be built, allowing for the energy of each gamma ray to be reconstructed.


\section{old}
Veritas

Veritas is a gamma-ray observatory.
This observatory consists of 4 telescopes, spread out over an area of 50 meters.
Each telescope consists of a 12m dish of 498 mirrors, with a customized 498 pixel single-photon timing camera at the end of a 7 meter boom.

Gamma rays are reconstructed and observed through a multi-stage process.
First, gamma rays are emitted from an astrophysical source, outside the solar system.
Then, these gamma rays travel to Earth and interact with molecules of Earth's atmosphere.
When they interact, they pair-produce into an electron-positron pair, where each has an energy of half the parent gamma ray.
The positron will then annihilate with an atmospheric electron, producing 2 more gamma rays, while the electron will shed some energy through ionization and by emitting chernekov photons.

This process will repeat over and over, creating a particle shower, until the electrons have ~80MeV of kinetic energy, at which point ionization will dominate the energy loss, and the shower will stop producing cherenkov light.

Veritas telescopes are designed to detect the cherenkov photons emitted from these particle showers, timing their arrival in each pixel to within 1 nanosecond.
The shower images produces by multiple telescopes can then be used to triangulate the original gamma ray's point of orgiin, whereas the size of the shower can be used to determine the gamma ray's energy.


Effective Area
For the calculation of the flux, the detection area must be found.
This depends primarily on each gamma ray's energy, as well as its detected position in the camera.

